---
title: "Final Project: Fraud"
author: "Group 9"
date: "2023-05-01"
output: html_document
---

```{r, include=FALSE} 
rm(list = ls())
library(tidyverse)
library(lubridate)
library(forcats)
library(RColorBrewer)
```

## Data Frame(s)

fraudTest = Entire Default Dataset

frauds = fraudTest where is_fraud = 1

not_frauds = fraudTest where is_fraud = 0

```{r}
fraudTest <- read.csv("class_data/fraudTest.csv")

frauds <- filter(fraudTest,is_fraud==1)

not_frauds <- filter(fraudTest,is_fraud==0)

str(fraudTest)
# str(frauds)
# str(not_frauds)
```

## Which jobs result in the highest fraudulent amount spent?

We created a new data frame called 'frauds_20', which shows the top 20 jobs, sorted by greatest 'amt_sum', or fraudulent amount sum. 

We started by grouping our 'frauds' data frame seen above by job, then using the summarize() function to create our amt_sum variable, then then the arrange() function to order our jobs from greatest to least. 


```{r setup, echo=FALSE}
frauds_20 = frauds %>% 
  group_by(job) %>% 
  summarize(amt_sum = sum(amt, na.rm = T)) %>%
  arrange(desc(amt_sum)) %>% 
  head(20)

frauds_20 = frauds_20 %>% 
  mutate(job2 = reorder(job, amt_sum, na.rm = TRUE, decreasing = FALSE))

frauds <- filter(fraudTest,is_fraud==1)
#Creates a dataset that is of cases of fraud
fraud_by_exp <- frauds %>% group_by(category) %>%
  summarise(Sum=sum(amt)/1000, Freq=n()) %>% arrange(desc(Sum))
#group the frauds by the expense
cc_by_exp <- fraudTest %>% group_by(category) %>%
  summarise(Total=sum(amt)/1000, Count=n()) %>% arrange(desc(Total))
#group the transaction by the expense
by_exp <- full_join(cc_by_exp,fraud_by_exp,join_by(category))
#joins the 2 exp datasets together
by_exp <- by_exp %>% mutate(Percent=Sum/Total/10,
                            Fraud_Percent=Freq/Count)
#creates the percent columns
ggplot(data=cc_by_exp)+
  geom_col(aes(x=reorder(category,Total), y=Total, fill=Total))+
  coord_flip()+
  labs(x="Expense",y="Money Spent (Thousands of Dollars)")+
  scale_fill_gradientn("Money Spent",colors =c("yellow","Red"))+
  ggtitle("Amount of CC Spending by Expense")
#plots the amount of cc spending by expense
#grocary leads this too
#shopping_pos is in close second
ggplot(data=fraud_by_exp)+
  geom_col(aes(x=reorder(category,Sum), y=Sum, fill=Sum))+
  coord_flip()+
  labs(x="Expense",y="Amount of Money (Thousands of Dollars)")+
  scale_fill_gradientn(colors =c("yellow","Red"))+
  ggtitle("Amount of Fraudulent Spending for Different Expenses")
#plot showing how much fraud each expense occurs
#Shopping_net leads them all

ggplot(data=by_exp)+
  geom_col(aes(x=reorder(category,Percent),y=Percent,fill=Percent))+
  coord_flip()+
  scale_fill_gradientn(colors =c("yellow","Red"))+
  labs(x="Expense",y="Percent")+
  ggtitle("Percent of Fraudulent Spending by Expense")
#plots how much cc spending on each expense is fraud
#shopping net is the leader of the group

ggplot(data=cc_by_exp)+
  geom_col(aes(x=reorder(category,Count), y=Count, fill=Count))+
  coord_flip()+
  labs(x="Expense",y="Frequency of Transactions")+
  scale_fill_gradientn("Times Used",colors =c("yellow","Red"))+
  ggtitle("Frequency of CC Purchases by Expense")
#plots how frequently cc is used to by something
#gas_transport leads this
#goscery_pos is in close second
ggplot(data=fraud_by_exp)+
  geom_col(aes(x=reorder(category,Freq), y=Freq, fill=Freq))+
  coord_flip()+
  labs(x="Expense",y="Frequency of Transactions")+
  scale_fill_gradientn(colors =c("yellow","Red"))+
  ggtitle("Frequency of Fraud for Expenses")
#plot showing how frequently each expense is frauded
#shopping_net leads this too, but grocery_pos is close
ggplot(data=by_exp)+
  geom_col(aes(x=reorder(category,Fraud_Percent),y=Fraud_Percent,fill=Fraud_Percent))+
  coord_flip()+
  scale_fill_gradientn("Fraud Percentage",colors =c("yellow","Red"))+
  labs(x="Expense",y="Percent")+
  ggtitle("Percent of Fraudulent Transactions by Expense")
#plots what percent of transaction are fraud
#one percent of 

fraud_by_state <- frauds %>% group_by(state) %>%
  summarise(Sum=sum(amt), Freq=n()) %>% arrange(state)
#groups all fradulent transactions by state
#creates state data, including total amount and frequency of fraud.
cc_by_state <- fraudTest %>% group_by(state) %>%
  summarise(Total=sum(amt)/1000, Count=n()) %>% arrange(state)
#groups all transactions by state
#creates state data, including total amount and frequency of fraud.
install.packages(c("maps", "mapdata","ggmap"))
library(ggmap)
library(maps)
library(mapdata)
#allows for mapping
States <- map_data("state")
#map data for states
By_state <- full_join(cc_by_state,fraud_by_state,by=join_by(state))
#joins my two state datasets together
By_state <- By_state[-c(1,8,9),]
#gets rid of AK, HI, DC as we don't have maping info
By_state <- By_state %>% mutate(Percent= Sum/Total/1000*100,
                                Fraud_Percent=(Freq/Count)*100)
#creates the percent and fraud percentage column 
By_state <- By_state %>% mutate(region=tolower(state.name[match(state,state.abb)]))
#create region to make it possible to join
States2 <- full_join(States,By_state,by=join_by(region))
#joins the two datasets together
States2["Sum"][is.na(States2["Sum"])] <- 0
States2["Percent"][is.na(States2["Percent"])] <- 0
States2["Freq"][is.na(States2["Freq"])] <- 0
States2["Total"][is.na(States2["Total"])] <- 0
States2["Count"][is.na(States2["Count"])] <- 0
States2["Fraud_Percent"][is.na(States2["Fraud_Percent"])] <- 0
#puts in zeros for any data we are missing

ggplot(data=cc_by_state)+
  geom_col(aes(x=reorder(state,Total), y=Total, fill=Total))+
  coord_flip()+
  labs(x="State",y="Amount of Money (Thousands of Dollars)")+
  scale_fill_gradientn(colors =c("yellow","Red"))+
  ggtitle("Amount of Credit Card Spending in US States")
#Texas spends the most money on credit cards, while NY and PA are second and third
#which makes sense as they lead in fraud
ggplot(data=States2,aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill=Total),color="black")+
  scale_fill_gradientn("CC Spending
  (1000's of $)",colors =c("Yellow","Red"))+
  ggtitle("Sum of CC Spending in the US")+
  labs(x="Longitude",y="Latitude")
#shows how much spending there is in each state
#NY, TX, PA is very high on the list
ggplot(data=fraud_by_state)+
  geom_col(aes(x=reorder(state,Sum), y=Sum, fill=Sum))+
  coord_flip()+
  labs(x="State",y="Amount of Fraud ($)")+
  scale_fill_gradientn("Sum of Fraud",colors=c("yellow","Red"))+
  ggtitle("Amount of Fraud in US States")
#creates graph for sum by state
#New York leads in that
ggplot(data=States2,aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill=Sum),color="black")+
  scale_fill_gradientn("Sum of Fraud",colors =c("Yellow","Red"))+
  ggtitle("Sum of Fraud in the US")+
  labs(x="Longitude",y="Latitude")
#plot to show how much fraud is in each state
#New York leads in fradulent spending
ggplot(data=States2,aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill=Percent),color="Black")+
  scale_fill_gradientn("Percent of Fraud",colors =c("Yellow","Red"))+
  ggtitle("Percent of Fraudulent Spending in the US")+
  labs(x="Longitude",y="Latitude")
#plots by state how much spending is Fraud
#Conneticut is leading the way
ggplot(data=cc_by_state)+
  geom_col(aes(x=reorder(state,Count), y=Count, fill=Count))+
  coord_flip()+
  labs(x="State",y="Fequency of Transactions")+
  scale_fill_gradientn(colors =c("yellow","Red"))+
  ggtitle("Use of Credit Cards in US States")
#Computes how often each state uses a credit card
#TX uses there credit card the most followed by NY, PA
ggplot(data=States2,aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill=Count),color="black")+
  scale_fill_gradientn("Frequency of 
  CC Use",colors =c("Yellow","Red"))+
  ggtitle("Frequency of CC Use in the US")+
  labs(x="Longitude",y="Latitude")
#plots how frequently each state uses a cc
ggplot(data=fraud_by_state)+
  geom_col(aes(x=reorder(state,Freq), y=Freq, fill=Freq))+
  coord_flip()+
  labs(x="State",y="Frequency of Fraud")+
  scale_fill_gradientn("Frequency",colors=c("yellow","Red"))+
  ggtitle("Frequency of Fraud in US States")
#Plots how frequently a state reports fraud
#NY has the most reported frauds
ggplot(data=States2,aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill=Freq),color="black")+
  scale_fill_gradientn("Frequency",colors =c("Yellow","Red"))+
  ggtitle("Amount of Fraudulent Transactions by State")+
  labs(x="Longitude",y="Latitude")
#plots for each state transactions are fraud
ggplot(data=States2,aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill=Fraud_Percent),color="black")+
  scale_fill_gradientn("Transaction
  Percentage",colors =c("Yellow","Red"))+
  ggtitle("Percent of Fraudulent Transactions by State")+
  labs(x="Longitude",y="Latitude")
#plots for each state what percent of transactions are fraud
usa <- map_data("usa")
#download map data for USA
gg1<- ggplot()+
  geom_polygon(data = usa, aes(x=long,y=lat, group=group),fill="white",color="black")+
  coord_fixed(1.3)
#creates function to plot USA data
gg1
lower_48_fraud <- frauds %>% filter(long>=-130)
lower_48_cc <- fraudTest %>% filter(long>=-130)
#gets rid of data outside the lower 48
gg1+
  geom_point(data=lower_48_cc, aes(x=long,y=lat),size=.1)+
  labs(x="Longitude",y="Latitude")+
  ggtitle("CC Transaction Location in US")
#plots where all transaction occur in US
gg1+
  geom_point(data=lower_48_fraud, aes(x=long,y=lat), color="black", size=.5)+
  labs(x="Longitude",y="Latitude")+
  ggtitle("CC Fraud Location in US")
#plots where all fraud occur in US
#like the transaction map just less dense
```

## Title

Description
-
-
-
-
-
-

```{r cars}
#code
```


```{r pressure, echo=FALSE}
# replace {r} with the above ^^ If you don't want the code to show
```

