**Analyzing Credit Card Fraud: A Case Study of Transactions from 2019-2020**
---
This research project aims to analyze a simulated credit card transaction dataset containing legitimate and fraudulent transactions from January 2019 to December 2020. The dataset includes credit card transactions made by 1000 customers with 800 merchants.
---

Here’s some code

```{r}
library(tidyverse)
```

```{r}
#reading in the csv file - 
data = read.csv(choose.files(), header = T)
```

```{r}
#DATA CLEANING - 
#checking for any missing values in the entire dataframe - 
sum(is.na(data))
#there are no missing values
#:)
```
```{r}
#1. converting trans_date_trans_time from a character variable to a datetime variable - for this we will be 
#creating a new variable tr_datetime that will contain said datetime variable.
data <- mutate(data,
             tr_datetime = dmy_hm(trans_date_trans_time))

#2. converting dob from a character variable to a date variable- for this we will be 
#creating a new variable DOB that will contain the date variable.
data <- mutate(data,
             DOB = dmy(dob))
```

How do demographics effect fraud credit card transactions? 

Age v/s Amt graph - 
```{r}
# Get current date
current_date <- as.Date(Sys.time())

#Calculate age based on birthdate - 
#convert DOB column to yyyy-mm-dd format-
data$DOB <- as.Date(data$DOB, format = "%Y-%m-%d")  

#creating a new variable called age that stores the ages of everyone- 
data$age <- as.numeric(format(current_date, "%Y")) - as.numeric(format(data$DOB, "%Y"))  

#Finding the minimum and the maximum age to see the range in the ages of the people:
min <- min(data$age)
min

max <- max(data$age)
max

#the min and max ages are 21 and 99 respectively. so, the breaks were made according 
#to that.
#categorizing the ages: 
data <- data %>% mutate(age_category = cut(age, breaks = c(20, 40, 60, 80, 100),
                                       labels = c("20-40", "40-60",
                                                  "60-80", "80-100")))


#graphing age vs amt graph - 
data <- filter(data,is_fraud==1)

fraud_by_age <- data %>% group_by(age_category) %>%
  summarise(Sum=sum(amt)/1000)

ggplot(data=fraud_by_age)+
  geom_point(aes(x= age_category, y=Sum, fill=Sum))+
  labs(x="age",y="Amount ($) in intervals of 1000") 

#from the graph, we can conclude that the people who have fallen into a 
#fraud are from the age group of 40-60 and they have lost approx. 
```
```{r}
##Analyzing how gender affects the amount of frauds
#Summarise and group by gender

data <- filter(data,is_fraud==1)

fraud_by_gender <- data %>% group_by(gender) %>%
  summarise(Sum=sum(amt/1000))

   str(fraud_by_gender)

#boxplot of fraud and Sum of amt
ggplot(data = data) +
  geom_boxplot(aes(x = gender, y = amt)) +
  labs(x = "Gender", y = "Amount", title = "Gender and Amount")

```

TRENDS IN FRAUDULENT TRANSACTIONS

cyclicity - how do fraudulent transactions distribute on the temporal spectrum?
is there an hourly, monthly, or seasonal trend?


```{r}
#HOURLY TREND:
data$hour <- factor(format(data$tr_datetime, "%H"), levels = sprintf("%02d", 0:23))

fraud_fraudTest <- data %>%
  filter(is_fraud == 1)

ggplot(data = fraud_fraudTest, aes(x = hour, fill = factor(is_fraud))) +
  geom_bar(position = "dodge", show.legend = FALSE) +
  scale_fill_manual(values = c("#E69F00"), name = "Type", labels = c("Fraud")) +
  labs(x = "Time (Hour) in a Day", y = "Count") +
  theme_minimal() +
  scale_x_discrete(limits = sprintf("%02d", seq(0, 23, 1)), labels = sprintf("%02d", seq(0, 23, 1)))

#The results from the graph are interesting! Most fraud cases happen from 10:00 p.m. 
#to 11 p.m. 
```

```{r}
#WEEKLY TREND:

data$day <- wday(data$tr_datetime)

data$day <- factor(data$day, levels = 1:7, labels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))

fraud_fraudTest <- data %>% 
  filter(is_fraud == 1)

ggplot(data = fraud_fraudTest, aes(x = day)) +
  geom_bar(fill = "#E69F00", show.legend = FALSE) +
  labs(x = "Day of Week", y = "Count", title = "Day of Week vs Fraud") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#Majority of the fraudulent transactions tend to happen on Mondays.
```

```{r}
#MONTHLY TRANSACTIONS:
data$month <- month(data$tr_datetime)

fraud_fraudTest <- data %>%
  filter(is_fraud == 1)

ggplot(fraud_fraudTest, aes(x = factor(month))) +
  geom_bar(fill = "#E69F00", show.legend = FALSE) +
  labs(x = "Month", y = "Count", title = "Month vs Fraud") +
  scale_x_discrete(labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#From the graph, we can tell that fraudulent transactions peak around March - May.
```

---
Next, let's take a look at what states are most affected by frauds.
---

```{r}
frauds <- filter(data,is_fraud==1)
#Creates a dataset that is of cases of fraud
fraud_by_state <- frauds %>% group_by(state) %>%
  summarise(Sum=sum(amt), Freq=n(), Avg=Sum/Freq) %>% arrange(state)
#creates state data, including total amount, freq, and average fraud.
ggplot(data=fraud_by_state)+
  geom_col(aes(x=reorder(state,Sum), y=Sum, fill=Sum))+
  coord_flip()+
  labs(x="State",y="Amount of Fraud ($)")+
  scale_fill_gradientn(colors=c("Blue","Red"))+
  ggtitle("Amount of Fraud in US States")

```
```{r}
cc_by_state <- data %>% group_by(state) %>%
  summarise(Total=sum(amt)/1000, Count=n()) %>% arrange(state)
```

```{r}
library(ggmap)
library(maps)
library(mapdata)
library(dplyr)
```

```{r}
#allows for mapping
States <- map_data("state")
#map data for states
By_state <- full_join(cc_by_state,fraud_by_state,by=join_by(state))
#joins my two state datasets together
By_state <- By_state[-c(1,8,9),]
#gets rid of AK, HI, DC as we don't have maping info
```

```{r}
By_state <- By_state %>% mutate(Percent= Sum/Total/1000*100,
                                Fraud_Percent=(Freq/Count)*100)
#creates the percent and fraud percentage column 
By_state <- By_state %>% mutate(region=tolower(state.name[match(state,state.abb)]))
#create region to make it possible to join
States2 <- full_join(States,By_state,by=join_by(region))
#joins the two datasets together
States2["Sum"][is.na(States2["Sum"])] <- 0
States2["Percent"][is.na(States2["Percent"])] <- 0
States2["Avg"][is.na(States2["Avg"])] <- 0
States2["Freq"][is.na(States2["Freq"])] <- 0
States2["Total"][is.na(States2["Total"])] <- 0
States2["Count"][is.na(States2["Count"])] <- 0
States2["Fraud_Percent"][is.na(States2["Fraud_Percent"])] <- 0
#puts in zeros for anydata we are missing
```

```{r}
usa <- map_data("usa")
#download map data for USA
gg1<- ggplot()+
  geom_polygon(data = usa, aes(x=long,y=lat, group=group),fill="white",color="black")+
  coord_fixed(1.3)
#creates function to plot USA data
gg1
```

```{r}
#plots by state how much spending is Fraud
ggplot(data=States2,aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill=Sum),color="black")+
  scale_fill_gradientn("Sum of Fraud",colors =c("Yellow","Red"))+
  ggtitle("Fraud in the US")+
  labs(x="Longitude",y="Latitude")
```

```{r}
lower_48_fraud <- frauds %>% filter(long>=-130)
lower_48_cc <- data %>% filter(long>=-130)
#gets rid of data outside the lower 48
gg1+
  geom_point(data=lower_48_fraud, aes(x=long,y=lat), color="black", size=.5)+
  labs(x="Longitude",y="Latitude")+
  ggtitle("CC Fraud Location in US")
```


